### 2011-02-07 Documentation/atropos/atropos.tex
Added some stuff.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@765 fd2da0a4-d968-48df-9cb5-d8c496287cbf

\usepackage{lscape}
\usepackage[round,semicolon,authoryear,]{natbib}

@@

\subsubsection{Supervised Semi-Interactive Segmentation}
Brain segmentation methods have relied on user interaction for many
years \citep{Lim1989,Julin1997,Freeborough1997a,Yushkevich2006}.
Atropos is capable of benefitting from user knowledge via an
initialization and optimization that depends upon a spatially
varying prior label image passed as input.  Rapid, sparse
labeling---with visualization provided by ITK-SNAP
(www.itksnap.org)---enables an interaction and execution processing
loop that can be critical to solving segmentation problems with
challenging clinical data in which automated approaches fail.  This
part of Atropos design is inspired by the interactive graph cuts
pioneered by \cite{Boykov2001} and which has spawned many follow-up
applications.  The Atropos prior label
image pre-specificies the segmentation results at a subset of the
spatial domain by fixing the priors and likelihood (and, thus, the
posterior) at a subset of $\mathcal{I}$ to be $1$ for the known label
and $0$ for each other label at the same site.  The user input
therefore not only initializes the optimization, but also gives boundary conditions
that influence the EM solution outside of the known sites.  While the
graph-based min-cut max-flow solution is globally optimal for two
labels, only locally optimal optimizers are available for 3 or more
classes.  Thus, in most practical applications, EM is a
reasonable and efficient alternative to Boykov's solution.
Furthermore, one may automate the initialization process.  
{\color{red}{We provide this capability to allow the user to implement
    an interactive editing and segmentation loop.  The user may
    run Atropos with sparse manual label guidance, evaluate the
    results, update the manual labels and repeat until achieving the desired
    outcome.  This processing loop be performed easily with, e.g., ITK-SNAP.}}

@@

model parameters, {\color{red}{Atropos is capable of updating the current label 

@@

\begin{figure}
\begin{center}
\begin{tabular}{cc}
\includegraphics[width=70mm]{r16slice.png} &
\includegraphics[width=70mm]{r16icm.png} \\
(a) & (b) \\
\end{tabular}
\end{center}
\caption{\baselineskip 12pt \small  (a) An adult brain image slice and (b) the corresponding ICM code image created using a $5\times5$ MRF neighborhood with a maximum ICM code label of `13'.  }
\label{fig:r16icm}
\end{figure}

@@

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
                           [initialMeshResolution,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output [correctedImage,<biasField>]
     -h 
     --help 
\end{lstlisting} 
\end{command}
\doublespacing
\subsubsection{Integrating N4 Bias Correction}
Assumptions about bias correction may be thought of as another prior
model.  As such, the typical segmentation processing pipeline begins with an intensity
normalization/bias correction step using a method such as the recently developed N4
algorithm \citep{Tustison2010}.  N4 builds upon the popular
non-parametric nonuniformity intensity normalization (N3) introduced in \cite{Sled1998}.  It was previously mentioned that
several methods proposed in the literature have taken an integrative
view of the segmentation problem by incorporating an intrinsic bias
correction step into the actual workflow.  The
advances introduced with N4 permit such an adaptive
integration with Atropos.  Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 \citep{Boyes2008}.  Thus, when performing 3-tissue segmentation, we
may opt to use, for instance, the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.  This is done by setting
the `{\ttfamily --weight-image}' option on the N4 command line call
(see Listing \ref{listing:n4}) to the posterior probability image
corresponding to the white matter produced as output in the Atropos
call, i.e. `{\ttfamily Atropos --output}'.  N4 was recently added  
to the Insight Toolkit repository%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1N4MRIBiasFieldCorrectionImageFilter.html
} 
where it is built and tested on
multiple platforms nightly.  The
evaluation section will illustrate inclusion of Atropos, N4 and ANTs
in a brain processing pipeline.

@@

\paragraph{Information Sharing Statement}
{\color{red}{Atropos software is available in ANTs
    \url{http://www.picsl.upenn.edu/ANTs} 
which depends on ITK \url{http://www.itk.org/Wiki/ITK/Git/Download}.
The data used in this work is available in the ANTs software
repository, BrainWeb 
\url{http://mouldy.bic.mni.mcgill.ca/BrainWeb/} and at \url{www.brain-development.org}.
We employed itk-SNAP for visualization \url{www.itksnap.org}.}

>>>>

\usepackage{lscape}
\usepackage[square,comma,numbers,sort&compress]{natbib}

@@

\subsubsection{Supervised Semi-Interactive Segmentation}
Brain segmentation methods have relied on user interaction for many
years \citep{Lim1989,Julin1997,Freeborough1997a,Yushkevich2006}.
Atropos is capable of benefitting from user knowledge via an
initialization and optimization that depends upon a spatially
varying prior label image passed as input.  Rapid, sparse
labeling---with visualization provided by ITK-SNAP
(www.itksnap.org)---enables an interaction and execution processing
loop that can be critical to solving segmentation problems with
challenging clinical data in which automated approaches fail.  This
part of Atropos design is inspired by the interactive graph cuts
pioneered by \cite{Boykov2001} and which has spawned many follow-up
applications.  The Atropos prior label
image pre-specificies the segmentation results at a subset of the
spatial domain by fixing the priors and likelihood (and, thus, the
posterior) at a subset of $\mathcal{I}$ to be $1$ for the known label
and $0$ for each other label at the same site.  The user input
therefore not only initializes the optimization, but also gives boundary conditions
that influence the EM solution outside of the known sites.  While the
graph-based min-cut max-flow solution is globally optimal for two
labels, only locally optimal optimizers are available for 3 or more
classes.  Thus, in most practical applications, EM is a
reasonable and efficient alternative to Boykov's solution.
Furthermore, one may automate the initialization process.  
{\color{blue}{We provide this capability to allow the user to implement
    an interactive editing and segmentation loop.  The user may
    run Atropos with sparse manual label guidance, evaluate the
    results, update the manual labels and repeat until achieving the desired
    outcome.  This processing loop be performed easily with, e.g., ITK-SNAP.}}

@@

model parameters, {\color{blue}{Atropos is capable of updating the current label 

@@

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=160mm]{ICMfigure.pdf}
%\includegraphics[width=70mm]{r16slice.png} &
%\includegraphics[width=70mm]{r16icm.png} \\
%(a) & (b) \\
\end{tabular}
\end{center}
\caption{\baselineskip 12pt \small {\color{blue}{An adult brain image slice is shown 
with its ICM code image corresponding to a $5\times5$ MRF neighborhood.  
To the right of the ICM code image, we focus on a single neighborhood
with a center voxel associated with the ICM code label of '10'.  Each 
center voxel in a specified neighborhood exhibits a unique ICM code
label which does not appear elsewhere in its neighborhood.  When performing
the segmentation labeling update for ICM, we iterate through the set
of ICM code labels and, for each code label, we iterate through the image
and update only those voxels associated with the current code label.}}
  }
\label{fig:r16icm}
\end{figure}

@@

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
                           [initialMeshResolution,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output [correctedImage,<biasField>]
     -h 
     --help 
\end{lstlisting} 
\end{command}
\doublespacing
\subsubsection{Integrating N4 Bias Correction}
Assumptions about bias correction may be thought of as another prior
model.  As such, the typical segmentation processing pipeline begins with an intensity
normalization/bias correction step using a method such as the recently developed N4
algorithm \citep{Tustison2010}.  {\color{blue}{N4 extends the popular
nonparametric nonuniform intensity normalization (N3) algorithm \citep{Sled1998}
in two principal ways:
\begin{itemize}
  \item We replace the least squares B-spline fitting with a parallelizable alternative 
  (which we also made publicly available in the Insight Toolkit)---
  the advantages being that 1) computation is much faster 
  and 2) smoothing is not susceptible to outliers as is characteristic with
  standard least squares fitting algorithms.
  \item The fitting algorithm permits a multi-resolution approach so whereas standard 
  N3 practice is to select a single resolution at which bias correction occurs, the N4
  framework permits a multi-resolution correction where a base resolution is chosen and
  correction can then occur at multiple resolution levels each resolution being twice 
  the resolution of the previous level.
\end{itemize}  
Specifically, with respect to segmentation, there exists a third advantage with N4 over
N3 in that the former permits the specification of a probabilistic mask as opposed 
to a binary mask.}}
%It was previously mentioned that
%several methods proposed in the literature have taken an integrative
%view of the segmentation problem by incorporating an intrinsic bias
%correction step into the actual workflow.  The
%advances introduced with N4 permit such an adaptive
%integration with Atropos.  
Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 \citep{Boyes2008}.  Thus, when performing 3-tissue segmentation, we
may opt to use, for instance, the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.  This is done by setting
the `{\ttfamily --weight-image}' option on the N4 command line call
(see Listing \ref{listing:n4}) to the posterior probability image
corresponding to the white matter produced as output in the Atropos
call, i.e. `{\ttfamily Atropos --output}'.  N4 was recently added  
to the Insight Toolkit repository%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1N4MRIBiasFieldCorrectionImageFilter.html
} 
where it is built and tested on
multiple platforms nightly.  The
evaluation section will illustrate inclusion of Atropos, N4 and ANTs
in a brain processing pipeline.

@@

\paragraph{Information Sharing Statement}
{\color{blue}{Atropos software is available in ANTs
    \url{http://www.picsl.upenn.edu/ANTs} 
which depends on ITK \url{http://www.itk.org/Wiki/ITK/Git/Download}.
The data used in this work is available in the ANTs software
repository, BrainWeb 
\url{http://mouldy.bic.mni.mcgill.ca/BrainWeb/} and at \url{www.brain-development.org}.
We employed itk-SNAP for visualization \url{www.itksnap.org}.}}
